### 2021-02-08 paper/model/initial_conditions.tex
Release v2.0.0.
Update model section.

Consider a situation where you want to start a simulation with the beginning set amidst the pandemic. It means that several thousands of individuals have already recovered from the disease, are infectious, symptomatic or in ICUs. Additionally, the sample of infectious people who will determine the course of the pandemic in the following periods is likely not representative of the whole population because of differences in behavior (number of contacts, assortativeness), past policies (school closures), etc.. The distribution of courses of diseases in the population at the begin of the simulation is called initial conditions.

To come up with realistic initial conditions, we match reported infections from official data to simulated individuals by available characteristics like age and geographic information. The matching must be done for each day of a longer time frame like a month to have individuals with possible health states. Then, health statuses evolve until the begin of the simulation period without simulating infections by contacts. It is also possible to correct reported infections for a reporting lag and scale them up to arrive at the true number of infections.

>>>>

Consider a situation where you want to start a simulation with the beginning set amidst
the pandemic. It means that several thousands of individuals should already have
recovered from the disease, be infectious, symptomatic or in intensive care at the start
of your simulation.
Additionally, the sample of infectious people who will determine the course of the
pandemic in the following periods is likely not representative of the whole population
because of differences in behavior (number of contacts, assortativeness), past policies
(school closures), etc..
The distribution of courses of diseases in the population at the begin of the simulation is
called initial conditions.

To come up with realistic initial conditions, we match reported infections from official
data to simulated individuals by available characteristics like age and geographic
information. The matching must be done for each day of a longer time frame like a month
to have individuals with possible health states. Then, health statuses evolve until the
begin of the simulation period without simulating infections by contacts.
We also correct reported infections for a reporting lag and scale them up to arrive
at the true number of infections.

### 2021-02-08 paper/model/matching.tex
Release v2.0.0.
Update model section.

The empirical data described above only allows to estimate the number of contacts each
person has. In order to simulate transmissions of Covid-19, the numbers of contacts has
to be translated into actual meetings between people. This is achieved by matching
algorithms:

@@

                if infection_takes_places(i, j):
                    # Update health statuses.

                remaining_contacts[i, contact_type] -= 1
                remaining_contacts[j, contact_type] -= 1
    \end{minted}
    \caption{Pseudo-code for the matching algorithm of non-recurrent contacts.}
\end{listing}

We first randomly draw a contact type and individual. For each contact of the drawn
contact type that person has, we first draw the group of the other person (first stage).
Next, we calculate the probability to be drawn for each member of the group, based on
the number of remaining contacts. I.e. people who have more remaining contacts are drawn
with a higher probability. This has to be re-calculated each time because with each
matched contact, the number of remaining contacts changes. We then draw the other
individual, determine whether an infection takes place and if so update the health
states of the newly infected person. Finally, we reduce the number of remaining contacts
of the two matched individuals by one.

>>>>

The empirical data described above only allows to estimate the number of contacts each
person has. In order to simulate transmissions of Covid-19, the numbers of contacts has
to be translated into actual meetings between people. This is achieved by our matching
algorithm:

@@

                if infection_takes_pinfectedlaces(i, j):
                    update_health_state_of_freshly_infected()

                remaining_contacts[i, contact_type] -= 1
                remaining_contacts[j, contact_type] -= 1
    \end{minted}
    \caption{Pseudo-code of the matching algorithm for non-recurrent contacts.}
\end{listing}

We first randomly draw a contact type and individual. For each contact of the drawn
contact type that person has, we first draw the group of the other person (first stage).
Next, we calculate the probability to be drawn for each member of the group, based on
the number of remaining contacts, i.e. people who have more remaining contacts are drawn
with a higher probability. This has to be re-calculated each time because with each
matched contact, the number of remaining contacts changes. We then draw the other
individual, determine whether an infection takes place and if so update the health
state of the newly infected person. Finally, we reduce the number of remaining contacts
of the two matched individuals by one.

### 2021-02-08 paper/model/number_of_contacts.tex
Updates and notes for more updates for the introduction.
Update model section.

Consider a hypothetical population of 1,000 individuals in which 50 were infected with a
novel infectious disease. From this data alone, it is impossible to say whether only
those 50 people had contact with an infectious person and the disease has an infection
probability of 1 in each contact or whether everyone met an infectious person but the
disease has an infection probability of only 5 percent per contact. SEIR models do not
even try to distinguish contact frequency from the infectiousness of each contact and
combine the two in one parameter that is not invariant to social distancing policies.

@@

The number of random and recurrent contacts at the workplace and at home is calibrated
with data provided by \citet{Mossong2008}. For details see
Section~\ref{sec:number_of_contacts}. In particular, we sample the number of contacts or
group sizes from empirical distributions that sometimes depend on age. It is also
possible to use economic or other behavioral models to predict the number of contacts.

Theoretically, each contact type can have its own infection probability. However, to
reduce the number of free parameters and thus avoid a potential over-fitting we impose
some constraints. For now, infection probabilities in schools, preschools and nurseries
are equal. Moreover, we restrict all work contacts to have the same infection
probability.\comment[id=T]{I have no idea about the current infection probabilities.
Does it need an update?}

>>>>

Consider a hypothetical population of 1,000 individuals in which 50 were infected with a
novel infectious disease. From this alone, it is impossible to say whether only
those 50 people had contact with an infectious person and the disease has an infection
probability of 1 in each contact or whether everyone met an infectious person but the
disease has an infection probability of only 5 percent per contact. SEIR models do not
distinguish contact frequency from the infectiousness of each contact and
combine the two in one parameter that is not invariant to social distancing policies.

@@

The number of random and recurrent contacts at the workplace, during leisure activities
and at home is calibrated with data provided by \citet{Mossong2008}. For details see
Section~\ref{sec:number_of_contacts}. In particular, we sample the number of contacts or
group sizes from empirical distributions that sometimes depend on age. It is also
possible to use economic or other behavioral models to predict the number of contacts.

Theoretically, each contact type can have its own infection probability. However, to
reduce the number of free parameters and thus avoid a potential over-fitting we impose
some constraints. For now, infection probabilities in schools, preschools and nurseries
are equal. Moreover, we restrict all work contacts and all non-work contacts to have
the same infection probability.

### 2021-02-08 paper/model/policies.tex
Enforced 88 character limit.
Update model section.

% importance of fine-grained contact types and having the entire population
This, in conjunction with our fine-grained contact types, allows us to easily implement a wide variety of policies. Allowing policies to depend on the health states of the entire population means that adaptive lockdowns where, for example, schools close when a certain threshold of infections is surpassed at the county level would be as simple as determining which counties are above the threshold and then setting all school contacts in these counties to zero.

@@

% contact tracing and how including the states means the sky is the limit
Some policies can be easily implemented if the background characteristics are suitably extended. For example, a schooling policy with split classes, where each half attends school every other week can be implemented by storing whether the child would attend in even or odd weeks in the background characteristics and then using that information in the policy function.

>>>>

This, in conjunction with our fine-grained contact types, allows us to easily implement
a wide variety of policies. Allowing policies to depend on the health states of the
entire population means that adaptive lockdowns where, for example, schools close when a
certain threshold of infections is surpassed at the county level would be as simple as
determining which counties are above the threshold and then setting all school contacts
in these counties to zero.

@@

Some policies can be easily implemented if the background characteristics are suitably
extended. For example, a schooling policy with A/B schooling, where each half attends
school every other week can be implemented by storing whether the child would attend in
even or odd weeks in the background characteristics and then using that information in
the policy function.

### 2021-02-08 paper/model/summary.tex
Release v2.0.0.
Update model section.

To alleviate these problems we propose a different model structure. Our model inherits
many features from agent-based simulation models but replaces the contacts between
moving particles by contacts between individuals who work, go to school, live in a
household and enjoy leisure activities. The structure of the model is depicted in
Figure~\ref{fig:model_graph}.

\begin{figure}[!tp]
    \centering
    \includegraphics[width=\textwidth]{../figures/model_detailed.png}
    \caption{Simplified graph of the model}
    \label{fig:model_graph}
\end{figure}

The background characteristics include age, county and occupation of each simulated
individual. Contact models are functions that map individual characteristics into a
predicted number of contacts. Currently we distinguish between eight types of contact
models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and
random work contacts, recurrent and random leisure contacts, and nursery, preschool, and
school contacts.

@@

Once a person is infected, the disease progresses in a fairly standard way which is also
depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild
symptoms are infectious for some time and recover eventually. Cases with severe symptoms
additionally require hospitalization and lead to either recovery or death.

There are two possible ways to enable testing for Covid-19 in the model. The first way
allows to specify the ratio of known to unknown infections for each day. A random sample
of all newly infected individuals will then receive a test result in the following days.
The second approach consists of three steps. First, individuals demand a test because
they, e.g., experience symptoms. Secondly, tests are allocated to individuals while
respecting governmental access restrictions to test. Thirdly, depending of the
capacities of laboratories, tests are processed for some time until the individual
receives her test result.

@@

Another key advantage of the model is that the number of contacts an individual has of
each contact type can be calibrated from publicly available data \citep{Mossong2008}.
This in turn allows us to estimate policy-invariant infection probabilities from time
series of infection and death rates using the method of simulated moments
\citep{McFadden1989}. Since the infection probabilities are time-invariant, data
collected since the beginning of the pandemic can be used for estimation. Moreover,
since we can model the testing strategies that were in place at each point in time, we
can correct the estimates for the fact that not all infections are observed.

>>>>

To predict the effects of a wide variety of fine-grained social distancing policies
we propose a different model structure.
Our model inherits many features from agent-based simulation models but replaces
the contacts between moving particles by contacts between individuals who work,
go to school, live in a household and enjoy leisure activities.
The structure of the model is depicted in Figure~\ref{fig:model_graph}.

\begin{figure}[!tp]
    \centering
    \includegraphics[width=\textwidth]{../figures/model_detailed.png}
    \caption{Simplified graph of the model.}
    \label{fig:model_graph}
\end{figure}

The background characteristics include age, county and occupation of each simulated
individual. Contact models are functions that map individual characteristics into a
predicted number of contacts. Currently, we distinguish between eight types of contact
models which are all listed in Figure~\ref{fig:model_graph}: households, recurrent and
random work contacts, recurrent and random leisure contacts, and nursery, preschool, and
school contacts.

@@

Once a person is infected, the disease progresses in a fairly standard way which is
depicted in Figure~\ref{fig:course_of_disease}. Asymptomatic cases and cases with mild
symptoms are infectious for some time and recover eventually. Cases with severe symptoms
additionally require hospitalization and lead to either recovery or death.

Our model also allows the evaluation of different testing strategies by having a detailed
three-step model of test demand, allocation and processing:
First, individuals demand a test because they, e.g., experience symptoms.
Secondly, tests are allocated to individuals while respecting governmental access
restrictions to test.
Thirdly, depending of the capacities of laboratories, tests are processed for some time
until the individual receives her test result.

@@

Another key advantage of the model is that the number of contacts an individual has of
each contact type can be calibrated from publicly available data \citep{Mossong2008}.
This in turn allows us to estimate policy-invariant infection probabilities from time
series of infection and death rates using the method of simulated moments
\citep{McFadden1989}. Since the infection probabilities are time-invariant, data
collected since the beginning of the pandemic can be used for estimation. Moreover,
since we model the testing strategies that were in place at each point in time, we
can correct the estimates for the fact that not all infections are observed.

### 2021-02-08 paper/model/testing.tex
Release v2.0.0.
Update model section.

The model offers two approaches to implement testing for Covid-19.

The first way can be described as an top-down approach. Using data on estimates for the ratio of known to all infections\footnotemark a random sample of individuals who is newly infected is sampled and assigned a test. The result of the test will be revealed after some duration which must take into account the time it takes for a person to develop symptoms, the availability of tests and time to process the tests.

\footnotetext{The Dunkelzifferradar project publishes daily estimates of the dark figure of infections under \url{https://covid19.dunkelzifferradar.de/}}

The second way is an bottom-up approach which consists of three phases. In the first phase, the demand for tests is modeled. Demand functions map from individual characteristics to a probability which is the probability for this individual to demand a test. There can be multiple demand functions where each function may describe a different channel. For example, individuals who experience symptoms or have a risk contact may ask for a test. Or, the ministry of education requires a negative test result from every teacher every second week. After the probabilities for each individual and every demand model are computed, individuals who demand a test as well as the channel is sampled.

@@

In the last and third phase, administered tests are determined to be processed. This step can become a bottleneck in the testing process if there are not enough laboratories or necessary resources available to evaluate all tests.

The strength of the first approach is its simplicity, but it relies on estimates for the ratio of known to unknown cases for which not much data exist in the absence of systematic and randomized testing. The second approach is much more detailed and in principle able to reflect individual responses and governmental restrictions on the access to tests. At the same time, individual behavior and restrictions are complex and the determinants of the demand, allocation and processing functions require a lot of calibration.

>>>>

We support to model testing as consisting of three stages.
Firstly, we model who demands a test. Demand functions map from individual
characteristics to a probability which is the probability for this individual to demand
a test. There can be multiple demand functions where each function may describe a
different channel. For example, individuals who experience symptoms or have a risk
contact may ask for a test. Or, the ministry of education requires a negative test
result from every teacher every second week. After the probabilities for each individual
and every demand model are computed, individuals who demand a test as well as the
channel is sampled.

@@

In the last and third phase, administered tests are processed.
This step can become a bottleneck in the testing process if there are not enough
laboratories or necessary resources available to evaluate all tests.

In our empirical estimation we use a very simplified testing model where the number
of tests to be distributed is calculated from estimates for the ratio of known to all
infections.\footnotemark
\footnotetext{The Dunkelzifferradar project publishes daily estimates of the dark figure
of infections under \url{https://covid19.dunkelzifferradar.de/}}
Using these estimates as well as data on the test distribution
over age groups by the RKI\footnotemark
\footnotetext{https://ars.rki.de/Content/COVID19/Main.aspx}
we allocate tests firstly among the symptomatic
and then randomly allocate tests to newly infected to fit the German test distribution.
